---
title: 'British Election Survey - Wave 9 - #Bregret over 65'
author: 'mz '
output:
  word_document: default
  pdf_document: default
pdf_document: default
html_document: default
---
  ```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
options(scipen = 9, digits = 4)
library(knitr)
library(dplyr)
library(tidyr)
```



```{r, echo = FALSE, message = FALSE, warning = FALSE}
# # Data was downloaded from # # Data was downloaded from http://www.britishelectionstudy.com/custom/uploads/2016/10/BES2015_W9_Panel_v1.0.zip
# # Then manually opened in Stata and saved as a csv file, so that only the required variables can be imported.
# bregret.all <- read.csv("data/BES2015_W9_Panel_v1.0/bregret.csv", nrows=1)

## looking for vairables:
# search.pattern <- "ageW7"
# names(bregret.all)[grep(search.pattern,names(bregret.all))]
# grep(search.pattern,names(bregret.all))
# 
# stencil <- c(rep("NULL",3879))
# stencil[c(18, #  = wt_full_W6
#           26, # = wt_full_W9
#           2218, # = regretsIHaveAFewW6
#           2636, # = euRefExpectationW7
#           3121, # = ageW7
#           3148, # = euRefExpectationW8
#           3464, # = econPersonalRetroW8
#           3509, # = ageW8
#           3531, # = euRefTurnoutRetroW9
#           3533, # = euRefVoteW9
#           3535, # = regretsIHaveAFewW9
#           3752, # = ageGroup
#           3758, # = ns_secW1W2W3W4W5
#           3759, # = ns_sec_analyticW1W2W3W4W5
#           3762, # = gender
#           3766, # = gor
#           3767 # = education
#           )] <- NA
# 
# bregret.raw <- read.csv("data/BES2015_W9_Panel_v1.0/bregret.csv", colClasses = stencil)
# write.table(bregret.raw, file = "data/bregret.raw.csv")

bregret.raw <- read.table( "../data/bregret.raw.csv")


bregret.raw %>%
  filter(!is.na(wt_full_W9)) %>%
  mutate(age = ifelse(!is.na(ageW8),ageW8, ageW7)) %>%
  filter(age > 65) -> over.65
```

## REGIONAL disaggregation of over 65s' regrets

From [wave nine of the British Election Study](http://www.britishelectionstudy.com/data-object/british-election-study-combined-wave-1-9-internet-panel/), there are a total of `r summarise(filter(over.65,!is.na(euRefVoteW9) & euRefVoteW9 != 9999 & !is.na(gor)), count = n())` respondents over 65, who voted in the referendum, (and remembered their vote, and had a valid record of the government office region they live in). The following is the regional disaggregation of their reported regrets. 

```{r, echo = FALSE, message = FALSE, warning = FALSE}


regrets.gor.w <- xtabs(wt_full_W9~ regretsIHaveAFewW9 + euRefVoteW9 + gor , data = over.65)

names(attributes(regrets.gor.w)$dimnames) <- c( "Regrets", "Reported vote", "GOR")

attributes(regrets.gor.w)$dimnames[[1]] <- c("Regrets - No", "Regrets - Yes", "Regrets - Not sure")
attributes(regrets.gor.w)$dimnames[[2]] <- c("Remain", "Leave", "Don't remember")


attributes(regrets.gor.w)$dimnames[[3]] <- c("North East", "North West", "Yorkshire and the Humber", "East Midlands", "West Midlands", "East of England", "London", "South East", "South West", "Wales", "Scotland")

regrets.gor.w %>% as.data.frame() %>%
filter(Reported.vote != "Don't remember") %>%
spread(Reported.vote, Freq) %>%
arrange(GOR, Regrets) %>%
select(GOR, Regrets, Remain, Leave) %>%
group_by(GOR) %>%
mutate(Remain = 100*Remain/sum(Remain),
Leave = 100*Leave/sum(Leave)) -> regrets.gor

N <- sum( xtabs(~ regretsIHaveAFewW9 + euRefVoteW9 + gor , data = over.65)[,1:2,1])
kable(regrets.gor[1:3,], digits = 2, caption = paste("(N =", N  , ")"))

N <- sum(xtabs(~ regretsIHaveAFewW9 + euRefVoteW9 + gor , data = over.65)[,1:2,2])
kable(regrets.gor[4:6,], digits = 2, caption = paste("(N =", N  , ")"))

N <- sum( xtabs(~ regretsIHaveAFewW9 + euRefVoteW9 + gor , data = over.65)[,1:2,3])
kable(regrets.gor[7:9,], digits = 2, caption = paste("(N =", N  , ")"))

N <- sum( xtabs(~ regretsIHaveAFewW9 + euRefVoteW9 + gor , data = over.65)[,1:2,4])
kable(regrets.gor[10:12,], digits = 2, caption = paste("(N =", N  , ")"))

N <- sum( xtabs(~ regretsIHaveAFewW9 + euRefVoteW9 + gor , data = over.65)[,1:2,5])
kable(regrets.gor[13:15,], digits = 2, caption = paste("(N =", N  , ")"))

N <- sum( xtabs(~ regretsIHaveAFewW9 + euRefVoteW9 + gor , data = over.65)[,1:2,6])
kable(regrets.gor[16:18,], digits = 2, caption = paste("(N =", N  , ")"))

N <- sum( xtabs(~ regretsIHaveAFewW9 + euRefVoteW9 + gor , data = over.65)[,1:2,7])
kable(regrets.gor[19:21,], digits = 2, caption = paste("(N =", N  , ")"))

N <- sum( xtabs(~ regretsIHaveAFewW9 + euRefVoteW9 + gor , data = over.65)[,1:2,8])
kable(regrets.gor[22:24,], digits = 2, caption = paste("(N =", N  , ")"))

N <- sum( xtabs(~ regretsIHaveAFewW9 + euRefVoteW9 + gor , data = over.65)[,1:2,9])
kable(regrets.gor[25:27,], digits = 2, caption = paste("(N =", N  , ")"))

N <- sum( xtabs(~ regretsIHaveAFewW9 + euRefVoteW9 + gor , data = over.65)[,1:2,10])
kable(regrets.gor[28:30,], digits = 2, caption = paste("(N =", N  , ")"))

N <- sum( xtabs(~ regretsIHaveAFewW9 + euRefVoteW9 + gor , data = over.65)[,1:2,11])
kable(regrets.gor[31:33,], digits = 2, caption = paste("(N =", N  , ")"))

```

The chart for the same data in Figure 1, with the widths of the bars representing the (weighted) size of each group of respondents. 


```{r fin2, fig.width=8,fig.height=9,echo=FALSE,message=FALSE,include=TRUE,fig.cap="Regrets by camp and region for respondents over 65"} 


regrets.gor %>% gather(vote, prop, 3:4) %>%
  unite(group , GOR, vote) %>%
  spread(group, prop)-> regrets.mat

par(mar = c(3, 12, 3, 1), xpd = TRUE)

bar.widths <- as.vector(apply(regrets.gor.w[,1:2,],2:3, sum)[2:1,])

barplot(as.matrix(regrets.mat[,2:23]),
        width = bar.widths,
        space = c(rep(c(1, 0.4), 11)),
        horiz =  TRUE, las = 2,
        legend.text = c("No regrets", "Yes regrets", "Not sure"),
        args.legend = list(x = 50, y = 40000))


```


