---
title: 'British Election Survey - Wave 9 - #Bregret'
author: 'mz '
output:
  pdf_document: default
  html_document: default
  word_document: default
---
```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
options(scipen = 9, digits = 4)
library(knitr)
library(dplyr)
library(tidyr)
```


The results of [wave nine of the British Election Study](http://www.britishelectionstudy.com/data-object/british-election-study-combined-wave-1-9-internet-panel/) have just recently been released (6.10.2016), the survey having been conducted in the days immediately after the EU referendum (24.6. to 4.7.). Over 30,000 respondents were asked about how they voted, and also whether they had any regrets about it in the immediate aftermath of the results being announced.


According to the [survey](https://github.com/majazaloznik/BrexitRegret), the proportion regretting their vote was significantly higher among leave voters (5.78 %) compared to remain voters (1.43 %).


```{r, echo = FALSE, message = FALSE, warning = FALSE}
# # Data was downloaded from http://www.britishelectionstudy.com/custom/uploads/2016/10/BES2015_W9_Panel_v1.0.zip
# # Then manually opened in Stata and saved as a csv file, so that only the required variables can be imported.
# bregret.all <- read.csv("data/BES2015_W9_Panel_v1.0/bregret.csv", nrows=1)
# stencil <- c(rep("NULL",3879))
# stencil[c(18, 26, 2218, 2636, 3148, 3464,3531,3533, 3535, 3758, 3759)] <- NA
# bregret.raw <- read.csv("data/BES2015_W9_Panel_v1.0/bregret.csv", colClasses = stencil)
# write.table(bregret.raw, file = "data/bregret.raw.csv")
bregret.raw <- read.table( "../data/bregret.raw.csv")
regrets <- xtabs(wt_full_W9~ regretsIHaveAFewW9 + euRefVoteW9 , data = bregret.raw)

names(attributes(regrets)$dimnames) <- c( "Regrets", "Reported vote")
colnames(regrets) <- c("Remain", "Leave", "Don't remember")
rownames(regrets) <- c("Regrets - No", "Regrets - Yes", "Regrets - Not sure")
regrets.prop <- apply(regrets, 2, function(x) x/sum(x))
N <- sum(xtabs(~ regretsIHaveAFewW9 + euRefVoteW9,data = bregret.raw))
kable(regrets.prop*100, digits = 2, caption = paste("Table 1: Percentage regretting their vote in each EU referendum camp (N =", N  , ")"))
```

Curiously regrets were even higher (6.29%) among the approximately 180 respondents who did not know how they had voted. But ignoring the latter, we can apply these (weighted) survey results to the actual referendum outcome: if these levels of regret are representative nationally and the referendum had been repeated only ten days or so later, what would the result have been? 


```{r, echo = FALSE, message = FALSE, warning = FALSE}
# # Manually get actual referendum result. 
# http://www.electoralcommission.org.uk/find-information-by-subject/elections-and-referendums/past-elections-and-referendums/eu-referendum/electorate-and-count-information
remain <- 16141241
leave <- 17410742


data.frame(regrets = c("no", "yes", "dont know"),
                       Remain = regrets.prop[,1]* remain,
                       Leave = regrets.prop[,2]* leave, row.names = NULL) %>%
  gather(vote, count, 2:3, convert = TRUE) %>%
  mutate(vote2 = c("Remain 2", "Leave 2", "Remain 2", "Leave 2", "Remain 2","Leave 2"))%>%
  gather(time, vote,c(2,4), convert = TRUE) %>%
  group_by(time, vote) %>%
  summarise(result = sum(count)) %>%
  ungroup() %>%
  group_by(time) %>%
  mutate(prop = result/sum(result)*100) ->
  results

kable(results[1:2,2:4], digits = 4, caption = "Table 2: Actual EU referendum results")

# results1 <- data.frame(regrets = c("0", "1", "0"),
#                        remain = regrets.prop[,1]* remain,
#                        leave = regrets.prop[,2]* leave, row.names = NULL) 
# results1 %>% 
#   group_by(regrets) %>% 
#   summarize(remain = sum(remain), leave = sum(leave))%>% 
#   gather(N1, Value, 2:3, convert = TRUE) %>%
#   mutate(N2 = c("remain2", "leave2",  "leave2", "remain2"),
#          edgecol = c("red", "red", "orange", "green")) %>%
#   arrange(N1, N2)->
#   edges
# 
# nodes <- data.frame(ID = unique(c(edges$N1, edges$N2)), x = c(1,1,2,2),
#                      col= c( "yellow", "blue", "yellow", "blue" ),
#                     stringsAsFactors= FALSE )
# 
# r <- makeRiver(nodes, as.data.frame(edges))
# plot(r, node_margin = 0)


```


Taking the numbers of voters [as they voted on June 23](http://www.electoralcommission.org.uk/find-information-by-subject/elections-and-referendums/past-elections-and-referendums/eu-referendum/electorate-and-count-information) (Table 2) we can see the difference between the two was `r results[1,3]-results[2,3]` voters. If we apply the percentages that regretted their vote according to the survey, this would mean around the same number of voters changing camps: Just over one million Brexiters would now vote to remain, while around a quarter of a million of those who had voted to remain, would now vote to leave (second row of Table 3). 

```{r, echo = FALSE, message = FALSE, warning = FALSE}


kable(data.frame(Remain = regrets.prop[,1]* remain,
                Leave = regrets.prop[,2]* leave),
      caption = "Table 3: Hypothetical numbers of voters regretting their decision"
)
```

If we are generous and leave those who were unsure or undecided about whether or not they regretted their decisions to stick with their original vote, this means the final result would indeed be reversed, but not by much: 

```{r, echo = FALSE, message = FALSE, warning = FALSE}


kable(results[3:4,2:4], digits = 2, caption = "Table 4: Hypothetical referendum results with regretters switching camp")

```

One should be careful in interpreting these results as indicative of what the result would be of a repeat referendum. 

For one, the regret data only refers to people who had *actually* voted. We do not know how strongly abstainers regret abstaining, or how that varies according to which camp they are in. Additionally, the regret data was collected in the 10 days after the referendum, and the situation may well be very different 4 months later, but we do not have data on that as reliable as the British Election Study is. 
```{r, echo = FALSE, message = FALSE, warning = FALSE}

comp <- cbind(elections2015 = prop.table(xtabs(wt_full_W6~ regretsIHaveAFewW6 , data = bregret.raw))*100,
      referendum2016 = prop.table(xtabs(wt_full_W9~ regretsIHaveAFewW9 , data = bregret.raw))*100)
rownames(comp) <- c("Regrets - No", "Regrets - Yes", "Regrets - Not sure")
options(scipen = 9, digits = 2)

```
Comparing the level of regret to the 2015 election we can see that the proportion without any regrets is indeed slightly smaller for the referendum (`r comp[1,2]` %) than the election (`r comp[1,1]` %). That is mainly due to the larger number of people *unsure* about whether they regretted their decision. In fact only `r comp[2,2]` % said they regretted their referendum decision compared to `r comp[2,1]` % in the election. 


```{r, echo = FALSE, message = FALSE, warning = FALSE}
options(scipen = 9, digits = 3)

kable(comp, digits = 2, caption = "Table 5: Comparing regret levels with 2015 elections",
      col.names = c("Elections 2015", "Referendum 2016"))
# pt <- cbind(elections2015 = xtabs(wt_full_W6~ regretsIHaveAFewW6 , data = bregret.raw),
  #    referendum2016 = xtabs(wt_full_W9~ regretsIHaveAFewW9 , data = bregret.raw))

# prop.test(c(pt[1,1], pt[1,2]), c(sum(pt[,1]), sum(pt[,2]))) 
purl("scripts/Regret.Rmd", output = "scripts/Bregret.R")
```

